---
title: "Data Manipulation"
output: github_document
---

This file is for doing data manipulation. 

```{r}
library(tidyverse)
```

Load the litters and pups datasets.

```{r}
litters_df =
  read_csv(
    "data/FAS_litters.csv",
    na = c("", "NA", "."))

litters_df = janitor::clean_names(litters_df)

pups_df =
  read_csv(
    "data/FAS_pups.csv",
    skip = 3,
    na = c("", "NA", "."))

pups_df = janitor::clean_names(pups_df)
```

## `select` some variables

```{r}
select(litters_df, group, litter_number)

select(litters_df, group:gd_of_birth)

select(litters_df, -group)

select(litters_df, starts_with("gd"))

select(litters_df, starts_with("gd"), group)

select(litters_df, LITTER = litter_number, everything())

select(litters_df, -starts_with("gd"))
```

a couple of alternatives

```{r}
rename(litters_df, GROUP = group)

relocate(litters_df, litter_number, gd0_weight)
```

In the pups data, select litter number, sex, and PD ears.

```{r}
select(pups_df, litter_number, sex, pd_ears)
```


## `filter` some rows

```{r}
filter(litters_df, gd_of_birth == 19)

filter(litters_df, pups_born_alive <= 5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)

filter(litters_df, pups_born_alive < 4 | pups_born_alive > 8 )

drop_na(litters_df, gd0_weight)
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2

```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```


## `mutate` to change some columns

let's add and change some columns

```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight, 
  wt_gain_squared = wt_gain ^ 2,
  group = str_to_lower(group))
```

In the pups dataset, create a variable that subtracts 7 from PD walk


```{r}
mutate(pups_df, pd_walk_minus_7 = pd_walk - 7)
```

## `arrange` the data


```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))
```

## Do multiple steps...

This is bad:

```{r}
litters_df = read_csv("data/FAS_litters.csv", na = c("", "NA", "."))
litters_rename_df = janitor::clean_names(litters_df)
litters_with_gd_df = select(litters_rename_df, group, starts_with("gd"))
litters_no_na_df = drop_na(litters_with_gd_df)
litters_wt_gain_df = mutate(litters_no_na_df, wt_gain = gd18_weight - gd0_weight)
```

This is worse:

```{r}
select(
  janitor::clean_names(
    read_csv("data/FAS_litters.csv", na = c("", "NA", "."))), 
  group, starts_with("gd"))
```

This is good:

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("", "NA", ".")) |> 
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight, 
    group = str_to_lower(group)
  )
```

Load pups, clean names, drop missing, keep litter number and pd variables, add pd walk -7

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv", skip = 3, na = c("", "NA", ".")) |> 
  janitor::clean_names() |> 
  drop_na() |> 
  select(litter_number, starts_with("pd")) |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
  )
```


sometime the first thing isn't a dataframe

```{r}
litters_df |> 
  filter(group %in% c("con7", "con8")) |> 
  lm(gd18_weight ~ gd0_weight, data = _)
```



`select` vs `pull`

```{r}
litters_df |> 
  select(group)

litters_df |> 
  filter(group %in% c("con7", "con8")) |> 
  pull(group)
```

never do `$`

```{r}
litters_df$group
```


pipes and viewing

```{r}
litters_df |> 
  filter(group %in% c("con7", "con8"))
```







